---
title: "Arrange partial dep. plots (GAM) 10K per year"
author: "Stenka Vulova"
date: '2024-05-15'
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
#library(raster)
library(terra)
library(rgdal)
library(lubridate)
#library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
#library(rgeos)

#library(grid)
#library(spatstat)
#library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

#library(ggspatial)
#library(geobuffer)
#library(rgeos)

library(tools)

library(data.table)

library(colorspace)

library(ggpointdensity)
library(mgcv)

```

# TCD

## Load dataset

50K dataset (10K from each year)

```{r load data }

comb_df_TCD = read.csv("D:/Stenka_Cliwac/Topic_1/05_RESULTS/20240426_red_dfs_depend_plots/10K_per_year/TCD_dep_df_50K.csv")

glimpse(comb_df_TCD)

#Rows: 50,000
#Columns: 3

```


## Plot 

```{r scatterplot }

# helps me decide how to set ylim
min(comb_df_TCD$SHAP_value) # -2.354232
max(comb_df_TCD$SHAP_value) # 2.226146

tic("ggplot")

dep_plot_TCD = ggplot(data = comb_df_TCD, aes(x = TCD, y = SHAP_value)) +
  geom_pointdensity(size = 1.4) +  # Scatter plot
  scale_color_viridis_c(name = "No. of neighbors")+
  geom_smooth(method = "loess", se = FALSE, color = "red", alpha = 0.5, size = 1.5) +  # Loess smoothing
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +  # Horizontal line at y=0
  labs(x = "Tree cover density (%)", y = "SHAP value") +  # Set x-axis label
  ylim(-2, 2) +  
  theme_bw()+ 
    theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)) 

dep_plot_TCD

ggsave('D:/Stenka_Cliwac/Topic_1/06_R_SCRIPTS/20240515_partial_dep_plots/dep_plot_allyears_TCD_50K_dens.png', plot=dep_plot_TCD, device='png', dpi=300, width=10, height= 5 )

toc()

tictoc::tic.log()

# ggplot: 50.31 sec elapsed

```

# broadleaf_perc

## Load dataset

50K dataset (10K from each year)

```{r load data }

comb_df_broadleaf_perc = read.csv("D:/Stenka_Cliwac/Topic_1/05_RESULTS/20240426_red_dfs_depend_plots/10K_per_year/broadleaf_perc_dep_df_50K.csv")

glimpse(comb_df_broadleaf_perc)

# comb_df$broadleaf_perc100 = comb_df$broadleaf_perc*100
comb_df_broadleaf_perc$broadleaf_perc100 = comb_df_broadleaf_perc$broadleaf_perc * 100

#Rows: 50,000
#Columns: 3

```


## Plot 

```{r scatterplot }

# helps me decide how to set ylim
min(comb_df_broadleaf_perc$SHAP_value) # -2.354232
max(comb_df_broadleaf_perc$SHAP_value) # 2.226146

tic("ggplot")

dep_plot_broadleaf_perc = ggplot(data = comb_df_broadleaf_perc, aes(x = broadleaf_perc100, y = SHAP_value)) +
  geom_pointdensity(size = 1.4) +  # Scatter plot
  scale_color_viridis_c(name = "No. of neighbors")+
  geom_smooth(method = "loess", se = FALSE, color = "red", alpha = 0.5, size = 1.5) +  # Loess smoothing
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +  # Horizontal line at y=0
  labs(x = "Broadleaf (%)", y = "SHAP value") +  # Set x-axis label
  ylim(-2, 2) +  
  theme_bw()+ 
    theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)) 

dep_plot_broadleaf_perc

ggsave('D:/Stenka_Cliwac/Topic_1/06_R_SCRIPTS/20240515_partial_dep_plots/dep_plot_allyears_broadleaf_perc_50K_dens.png', plot=dep_plot_broadleaf_perc, device='png', dpi=300, width=10, height= 5 )

toc()

tictoc::tic.log()

# ggplot: 50.31 sec elapsed

```

# canopyheight

## Load dataset

50K dataset (10K from each year)

```{r load data }

comb_df_canopyheight = read.csv("D:/Stenka_Cliwac/Topic_1/05_RESULTS/20240426_red_dfs_depend_plots/10K_per_year/canopyheight_dep_df_50K.csv")

glimpse(comb_df_canopyheight)

#Rows: 50,000
#Columns: 3

```


## Plot 

```{r scatterplot }

# helps me decide how to set ylim
min(comb_df_canopyheight$SHAP_value) # -2.354232
max(comb_df_canopyheight$SHAP_value) # 2.226146

tic("ggplot")

dep_plot_canopyheight = ggplot(data = comb_df_canopyheight, aes(x = canopyheight, y = SHAP_value)) +
  geom_pointdensity(size = 1.4) +  # Scatter plot
  scale_color_viridis_c(name = "No. of neighbors")+
  geom_smooth(method = "loess", se = FALSE, color = "red", alpha = 0.5, size = 1.5) +  # Loess smoothing
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +  # Horizontal line at y=0
  labs(x = "Canopy height (m)", y = "SHAP value") +  # Set x-axis label
  ylim(-2, 2) +  
  theme_bw()+ 
    theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)) 

dep_plot_canopyheight

ggsave('D:/Stenka_Cliwac/Topic_1/06_R_SCRIPTS/20240515_partial_dep_plots/dep_plot_allyears_canopyheight_50K_dens.png', plot=dep_plot_canopyheight, device='png', dpi=300, width=10, height= 5 )

toc()

tictoc::tic.log()

# ggplot: 50.31 sec elapsed

```

# SPEI_sameyear

## Load dataset

50K dataset (10K from each year)

```{r load data }

comb_df_SPEI_sameyear = read.csv("D:/Stenka_Cliwac/Topic_1/05_RESULTS/20240426_red_dfs_depend_plots/10K_per_year/SPEI_sameyear_dep_df_50K.csv")

glimpse(comb_df_SPEI_sameyear)

#Rows: 50,000
#Columns: 3

```


## Plot 

```{r scatterplot }

# helps me decide how to set ylim
min(comb_df_SPEI_sameyear$SHAP_value) # -2.354232
max(comb_df_SPEI_sameyear$SHAP_value) # 2.226146

tic("ggplot")

dep_plot_SPEI_sameyear = ggplot(data = comb_df_SPEI_sameyear, aes(x = SPEI_sameyear, y = SHAP_value)) +
  geom_pointdensity(size = 1.4) +  # Scatter plot
  scale_color_viridis_c(name = "No. of neighbors")+
  geom_smooth(method = "loess", se = FALSE, color = "red", alpha = 0.5, size = 1.5) +  # Loess smoothing
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +  # Horizontal line at y=0
  labs(x = "SPEI (same year) (-)", y = "SHAP value") +  # Set x-axis label
  ylim(-2, 2) +  
  theme_bw()+ 
    theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)) 

dep_plot_SPEI_sameyear

ggsave('D:/Stenka_Cliwac/Topic_1/06_R_SCRIPTS/20240515_partial_dep_plots/dep_plot_allyears_SPEI_sameyear_50K_dens.png', plot=dep_plot_SPEI_sameyear, device='png', dpi=300, width=10, height= 5 )

toc()

tictoc::tic.log()

# ggplot: 50.31 sec elapsed

```


# ET_1year

## Load dataset

50K dataset (10K from each year)

```{r load data }

comb_df_ET_1year = read.csv("D:/Stenka_Cliwac/Topic_1/05_RESULTS/20240426_red_dfs_depend_plots/10K_per_year/ET_1year_dep_df_50K.csv")

glimpse(comb_df_ET_1year)

#Rows: 50,000
#Columns: 3

```


## Plot 

```{r scatterplot }

# helps me decide how to set ylim
min(comb_df_ET_1year$SHAP_value) # -2.354232
max(comb_df_ET_1year$SHAP_value) # 2.226146

tic("ggplot")

dep_plot_ET_1year = ggplot(data = comb_df_ET_1year, aes(x = ET_1year, y = SHAP_value)) +
  geom_pointdensity(size = 1.4) +  # Scatter plot
  scale_color_viridis_c(name = "No. of neighbors")+
  geom_smooth(method = "loess", se = FALSE, color = "red", alpha = 0.5, size = 1.5) +  # Loess smoothing
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +  # Horizontal line at y=0
  labs(x = "ET, annual sum (1 year ago) (mm)", y = "SHAP value") +  # Set x-axis label
  ylim(-2, 2) +  
  theme_bw()+ 
    theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)) 

dep_plot_ET_1year

ggsave('D:/Stenka_Cliwac/Topic_1/06_R_SCRIPTS/20240515_partial_dep_plots/dep_plot_allyears_ET_1year_50K_dens.png', plot=dep_plot_ET_1year, device='png', dpi=300, width=10, height= 5 )

toc()

tictoc::tic.log()

# ggplot: 50.31 sec elapsed

```

# forest_prox

## Load dataset

50K dataset (10K from each year)

```{r load data }

comb_df_forest_prox = read.csv("D:/Stenka_Cliwac/Topic_1/05_RESULTS/20240426_red_dfs_depend_plots/10K_per_year/forest_prox_dep_df_50K.csv")

glimpse(comb_df_forest_prox)

#Rows: 50,000
#Columns: 3

```


## Plot 

```{r scatterplot }

# helps me decide how to set ylim
min(comb_df_forest_prox$SHAP_value) # -2.354232
max(comb_df_forest_prox$SHAP_value) # 2.226146

tic("ggplot")

dep_plot_forest_prox = ggplot(data = comb_df_forest_prox, aes(x = forest_prox, y = SHAP_value)) +
  geom_pointdensity(size = 1.4) +  # Scatter plot
  scale_color_viridis_c(name = "No. of neighbors")+
  geom_smooth(method = "loess", se = FALSE, color = "red", alpha = 0.5, size = 1.5) +  # Loess smoothing
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", size = 0.8) +  # Horizontal line at y=0
  labs(x = "Distance to forest edge (m)", y = "SHAP value") +  # Set x-axis label
  ylim(-2, 2) +  
  theme_bw()+ 
    theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)) 

dep_plot_forest_prox

ggsave('D:/Stenka_Cliwac/Topic_1/06_R_SCRIPTS/20240515_partial_dep_plots/dep_plot_allyears_forest_prox_50K_dens.png', plot=dep_plot_forest_prox, device='png', dpi=300, width=10, height= 5 )

toc()

tictoc::tic.log()

# ggplot: 50.31 sec elapsed

```


# Join the tables together 

```{r join the tables together }

glimpse(comb_df_TCD)
colnames(comb_df_TCD)[1] <- "predictor_value"
comb_df_TCD$Predictor = 'TCD'

colnames(comb_df_broadleaf_perc)[1] <- "predictor_value"
comb_df_broadleaf_perc$predictor_value = comb_df_broadleaf_perc$predictor_value *100
comb_df_broadleaf_perc$Predictor = 'broadleaf_perc'
comb_df_broadleaf_perc$broadleaf_perc100 = NULL

colnames(comb_df_canopyheight)[1] <- "predictor_value"
comb_df_canopyheight$Predictor = 'canopyheight'

colnames(comb_df_SPEI_sameyear)[1] <- "predictor_value"
comb_df_SPEI_sameyear$Predictor = 'SPEI_sameyear'

colnames(comb_df_ET_1year)[1] <- "predictor_value"
comb_df_ET_1year$Predictor = 'ET_1year'

colnames(comb_df_forest_prox)[1] <- "predictor_value"
comb_df_forest_prox$Predictor = 'forest_prox'

df_allpreds = rbind(comb_df_broadleaf_perc, comb_df_canopyheight, comb_df_ET_1year, comb_df_forest_prox, comb_df_SPEI_sameyear, comb_df_TCD)

glimpse(df_allpreds)
# Rows: 300,000
# Columns: 4

#6 *50000

```

# Facet_wrap plot

```{r strip labels }

my_strip_labels <- as_labeller(c(
  "TCD" = "Tree cover density (%)",
  "broadleaf_perc" = "Broadleaf (%)",
  "canopyheight" = "Canopy height (m)",
   "SPEI_sameyear" = "SPEI (same year) (-)",
  "ET_1year" = "ET, annual sum (1 year ago) (mm)",
  "forest_prox" = "Distance to forest edge (m)"
))

```

```{r ggplot facet_wrap }

tic("ggplot facet")

dep_plot_facet = ggplot(data = df_allpreds, aes(x = predictor_value, y = SHAP_value)) +
  geom_pointdensity(size = 1.4) +  # Scatter plot
  scale_color_viridis_c(name = "No. of neighbors")+
  geom_smooth(method = "gam", se = FALSE, color = "red", alpha = 0.5, size = 1.3) +  
  geom_hline(yintercept = 0, linetype = "dashed", color = "orange2", size = 1) +  # Horizontal line at y=0
  labs(x = NULL, y = "SHAP value") +  # Set x-axis label
  ylim(-2, 2) +  
  theme_bw()+ 
    facet_wrap(
    facets = ~Predictor, scales = "free_x", 
    labeller = my_strip_labels,  # add labels
    strip.position = "bottom") +
  theme(
    strip.placement = "outside",
    strip.background = element_blank()
  )+
    theme(axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title = element_text(size = 20),
        strip.text = element_text(size = 20),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 16)) 

dep_plot_facet

ggsave('D:/Stenka_Cliwac/Topic_1/06_R_SCRIPTS/20240522_partial_dep_plots_GAM/dep_plot_allpredictors_50K_GAM.png', plot=dep_plot_facet, device='png', dpi=300, width=17, height= 10 )

toc()

tictoc::tic.log()

# ggplot facet: 72.67 sec elapsed

```


