---
title: "SHAP figure for SPEI_sameyear and 2018"
author: "Stenka Vulova"
date: '2024-04-16'
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

library(ggspatial)
#library(geobuffer)
library(rgeos)

library(tools)
library(tibble)

```

# Import Brandenburg and Berlin border

Keep Berlin in the plots :) 

```{r import border }

BB_border = readOGR(dsn = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border",
                        layer = "BB_border_merged")

crs(BB_border)
# CRS arguments: +proj=longlat +datum=WGS84 +no_defs 

plot(BB_border)

# +proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs
#BB_LAEA = sp::spTransform(BB_border, CRSobj = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

```

# Load training df

You don't need to run this part. I just used it to get which column number each predictor has.
You need to add 1 to the predictor column number (see below); for example `19. SPEI_sameyear` in the `trainX_df` is column 20 in the SHAP dfs (because first column in SHAP dfs is the modelled class).

This was used for training. I will use this df to infer the column # of the predictor I want to plot. 

```{r load df }

tic("load train df")

trainX_df = read.csv("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Dfs_for_SHAP_mapping/X_all_2018.csv")

toc()

tic.clearlog()

#load train df: 816.68 sec elapsed

glimpse(trainX_df)

summary(trainX_df)

#Rows: 10,406,171
#Columns: 41

# get the column names 

colnames_trainX = colnames(trainX_df)

write.csv(colnames_trainX, )

colnames_trainX[41]
# "SWI_15cm_1year"

# Create a list of column names with their corresponding column numbers
paste0(seq_along(colnames_trainX), ". ", colnames_trainX)

#  [1] "1. ag_prox"               "2. aspect"                "3. bdod"                  "4. broadleaf_perc"        "5. SPEI_sameyear"          "6. cec"                  
#  [7] "7. cfvo"                  "8. clay"                  "9. elevation"             "10. forest_prox"          "11. AWC"                  "12. nitrogen"            
# [13] "13. ocd"                  "14. ocs"                  "15. sand"                 "16. silt"                 "17. slope"                "18. soc"                 
# [19] "19. SPEI_sameyear"                  "20. TWI"                  "21. water_prox"           "22. airtemp_sameyear"     "23. ET_sameyear"          "24. rad_direct_sameyear" 
# [25] "25. SMI_sameyear"         "26. SPEI_sameyear"        "27. SSM_sameyear"         "28. SWI_15cm_sameyear"    "29. groundwater_sameyear" "30. gw_trend_sameyear"   
# [31] "31. LST_sameyear"         "32. airtemp_1year"        "33. ET_1year"             "34. groundwater_1year"    "35. gw_trend_1year"       "36. LST_1year"           
# [37] "37. rad_direct_1year"     "38. SMI_1year"            "39. SPEI_1year"           "40. SSM_1year"            "41. SWI_15cm_1year" 

```

# SHAP dataframe

```{r import the df }

tic("load SHAP df")

# # Load only columns of the SHAP dataframe
# cols_SHAPdf = read.table("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240415_SHAP.df_NEW/colnames_SHAPdf.csv", sep = ",", header = FALSE)
# 
# # Select the top highest ranked predictors based on the "all years" SHAP summary plot. 
# top_pred_names = c("TCD", "broadleaf_perc", "SPEI_sameyear", "SPEI_sameyear", "ET_1year")
# 
# # Match the names in top_pred_names with the values in the "V1" column of cols_SHAPdf
# matched_rows <- match(top_pred_names, cols_SHAPdf$V1, nomatch = 0)

shap_df_path = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2018_SHAP.csv"

# Use fread() to read only the first column
# first column is the modeled class

# SPEI_sameyear is column 26 originally, so it is column 27 here
# because first column here is the predicted classes, then the SHAP values for the predictors

shap_df <- data.table::fread(shap_df_path, select = 27, header = TRUE)

toc()

tic.clearlog()

glimpse(shap_df)
# Rows: 7,708
# Columns: 25

#just column 1 
#load SHAP df: 82.26 sec sec elapsed
#Rows: 10,406,171
#Columns: 1

# I expect there to be 42 columns in the SHAP df. 
# 1st column is the class and 41 cols for predictors' SHAP values

#rror in data.table::fread(shap_df_path, select = 43, header = TRUE) : 
#  Column number 43 (select[1]) is too large for this table, which only has 42 columns.

# yes, it has 42 cols

```

# Coordinates dataframe 

I am loading the coordinates. 

```{r load coords }

#true_class <- mod_df[, (names(mod_df) %in% c("x", "y", "coord_code", "true_class", "NDMI_anomaly"))]

true_class_path = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230927_modeling_df/true_class/true_class_2018.csv"

# first two columns are x and y 

coords_df <- data.table::fread(true_class_path, select = c(1,2), header = TRUE)

glimpse(coords_df)

# Rows: 10,406,171
# Columns: 2
# x
# y

```
# Make into one df 

x, y, and SHAP value of a certain predictor 

```{r new df with coords and SHap values }

shap_df_coords = cbind(coords_df, shap_df)

glimpse(shap_df_coords)

# Rows: 10,432,546
#Columns: 3
# x
# y
# `18`

# rename third column to SHAP value

names(shap_df_coords)[3] <- "SHAP_value"

```

# Convert to raster 

## Load static data

Static predictors co-registered to NDMI. 
This is just to get the CRS from it. No need to run this.

```{r Load static data }

#examp_rast = raster::raster("D:/Stenka_Cliwac/Topic_1/04_PROCESSED_DATA/20230926_data_proj/Static_data_proj/broadleaf_proj.tif")

#raster::plot(examp_rast, col=colorRampPalette(c("#460000", "#930c0c", "#ff2100", "#ff980c",
#                                 "#fcf599", "#00ff9b", "#39b5ff","#002ae7",
#                                "#020d47"))(255))

#examp_rast

# crs        : +proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs

```

## Raster conversion

```{r convert to raster }

SHAP_rast = terra::rast(shap_df_coords, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

# the reason I converted to lonlat is to include the scalebar (it didnt work with the other projection)
SHAP_rast_lonlat = terra::project(x = SHAP_rast, y = "+proj=longlat +datum=WGS84 +no_defs")

# convert to raster::raster for plotting
SHAP_r = raster::raster(SHAP_rast_lonlat)

raster::plot(SHAP_r, colNA = "lightgray", col=colorRampPalette(c("#460000", "#930c0c", "#ff2100", "#ff980c",
                                 "#fcf599", "#00ff9b", "#39b5ff","#002ae7",
                                "#020d47"))(255),
             legend.args = list(text = "SHAP value", line = 0.5, cex = 1, adj = 0.25))

minValue(SHAP_r) # -0.5437935
maxValue(SHAP_r) # 0.6043676


```

# Final plot 

The one I am saving. 

```{r plot it}


png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240422_SHAP_predictor_maps/20240423_SPEI_sameyear_SHAP_plots/SPEI_sameyear_SHAP_map_2018.png", width = 8, height =7, units = "in", res = 300)

raster::plot(SHAP_r, col = colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255),
             legend.args = list(text = "SHAP value", line = 0.5, cex = 1, adj = 0.25),
   #               xlim = c(11.26573, 14.7657), # Adjust the x-axis limits accordingly
   #  ylim = c(51.35902, 53.5587), # Adjust the y-axis limits accordingly
 #                 zlim = c(-1.771957, 2.35251),
             legend = TRUE, axes = FALSE, box = FALSE)


# only the legend
#plot(SHAP_r, col = colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255) ,
#     legend.only = TRUE, horizontal = FALSE , smallplot = c(0.87, 0.9, 0.17, 0.86),
 #                      zlim = c(-1.771957, 2.35251),
 #    axis.args = list(cex.axis = 1.8), 
#	legend.args = list(text = "SHAP value", line = 0.5, cex = 1, adj = 0.25))

# add scalebar 
raster::scalebar(d = 50, # distance in km,
                 xy = c(11.5, 51.5),
                 type = "bar",
                 divs = 2,
                 below = "km",
                 lonlat = TRUE,
                adj = c(0.5, -1.15),
                 cex = 1.2)

# Berlin Brandenburg border
raster::plot(BB_border, add = TRUE)


dev.off()

```


