---
title: "SPEI_annual_maps"
author: "Katharina Horn"
date: "2024-05-14"
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

library(ggspatial)
#library(geobuffer)
library(rgeos)

library(tools)
library(tibble)

```


# Import Brandenburg and Berlin border

Keeping Berlin in the plots.

``` {r Load Berlin-Brandenburg border}

bb <- terra::vect("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp") # for cropping of data frames

# Load as sf object
bb_st <- st_read("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp") # for mapping

```


# SHAP dataframe

```{r import the df }

# tic("load SHAP df")
# 
# # Define path
# shap_df_path_2018 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2018_SHAP.csv"
# shap_df_path_2019 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2019_SHAP.csv"
# shap_df_path_2020 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2020_SHAP.csv"
# shap_df_path_2021 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2021_SHAP.csv"
# shap_df_path_2022 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2022_SHAP.csv"

# Read only the header of the one of the files using read.table
# header <- read.table("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2018_SHAP.csv", header = TRUE, nrows = 1, sep = ",")
# header
# Second row is the first predictor (X0.1) followed by the other predictors (X1, X2, ...)

# # Use fread() to read only the first column
# # first column is the modeled class
# 
# # SPEI_sameyear is column 26 originally, so it is column 27 here
# # because first column here is the predicted classes, then the SHAP values for the predictors
# 
# shap_df_2018 <- data.table::fread(shap_df_path_2018, select = 27, header = TRUE)
# shap_df_2019 <- data.table::fread(shap_df_path_2019, select = 27, header = TRUE)
# shap_df_2020 <- data.table::fread(shap_df_path_2020, select = 27, header = TRUE)
# shap_df_2021 <- data.table::fread(shap_df_path_2021, select = 27, header = TRUE)
# shap_df_2022 <- data.table::fread(shap_df_path_2022, select = 27, header = TRUE)
# 
# toc()
# 
# tic.clearlog()


# Load dataframe that contains SPEI SHAP values of all years
shap_df_path = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/SPEI_sameyear_mean_SHAP_df.csv"

# column names: "x"         "y"         "2018"      "2019"      "2020"      "2021"      "2022"      "mean_SHAP"

# Use fread() to read specific columns
# I need all years and the coordinates x & y

shap_df <- data.table::fread(shap_df_path, select = c(1,2,3,4,5,6,7), header = TRUE)

glimpse(shap_df)


```
```{r Create subsets for each year}

# Subset for year 2018
subset_2018 <- shap_df[, c("x", "y", "2018")]
subset_2019 <- shap_df[, c("x", "y", "2019")]
subset_2020 <- shap_df[, c("x", "y", "2020")]
subset_2021 <- shap_df[, c("x", "y", "2021")]
subset_2022 <- shap_df[, c("x", "y", "2022")]


# Convert to data frame
SPEI_2018 <- data.frame(subset_2018)
SPEI_2019 <- data.frame(subset_2019)
SPEI_2020 <- data.frame(subset_2020)
SPEI_2021 <- data.frame(subset_2021)
SPEI_2022 <- data.frame(subset_2022)

```


```{r Define color scale}

colors <- colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255)

```

# Plot mean SHAP of all five predictors

```{r Plot mean SHAP of all five predictors separately}

color_limits_SPEI <- c(-0.6, 0.6)
# Plot SPEI_sameyear (2018)

SPEI_2018_plot <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = X2018), data = SPEI_2018) +             # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]),
                       na.value="white", oob = scales::squish) +                                # set legend colours
  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal", legend.position = "bottom") +
  theme_void() +
  theme(legend.position = "bottom")                                       # no theme

#SPEI_2018_plot


# Plot SPEI_sameyear (2019)

SPEI_2019_plot <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = X2019), data = SPEI_2019) +             # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]),
                       na.value="white", oob = scales::squish) +                                # set legend colours
  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal", legend.position = "bottom") +
  theme_void() +
  theme(legend.position = "bottom")                                       # no theme

#SPEI_2019_plot

# Plot SPEI_sameyear (2020)

SPEI_2020_plot <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = X2020), data = SPEI_2020) +             # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]),
                       na.value="white", oob = scales::squish) +                                # set legend colours
  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal", legend.position = "bottom") +
  theme_void() +
  theme(legend.position = "bottom")                                       # no theme

#SPEI_2020_plot

# Plot SPEI_sameyear (2021)

SPEI_2021_plot <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = X2021), data = SPEI_2021) +            # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]),
                       na.value="white", oob = scales::squish) +                               # set legend colours
  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal", legend.position = "bottom") +
  theme_void() +
  theme(legend.position = "bottom")                                      # no theme

#SPEI_2021_plot

# Plot SPEI_sameyear (2022)

SPEI_2022_plot <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = X2022), data = SPEI_2022) +            # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]),
                       na.value="white", oob = scales::squish) +                               # set legend colours
  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal", legend.position = "bottom") +
  theme_void() +
  theme(legend.position = "bottom")                                      # no theme

#SPEI_2022_plot


```


# Arrange and export plots

In Inkscape, I will add the label headings (aligning did not work out well within ggarrange() and I will add the scales as well. 

```{r Arrange resulting plots}

png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240422_SHAP_predictor_maps/20240423_SPEI_sameyear_SHAP_plots/SPEI_SHAP_allyears.png", 
    width = 17.5, height = 12, 
    units = "cm", res = 300)

SPEI_all_years <- ggarrange(SPEI_2018_plot, SPEI_2019_plot, SPEI_2020_plot, SPEI_2021_plot, SPEI_2022_plot,
                            labels = c("2018", "2019", "2020", "2021", "2022"),
                            ncol = 3, nrow = 2, 
                            legend = "none")

SPEI_all_years

dev.off()

```

# Get legend of one of the plots

I can extract a legend from one of the plots since all of them are based on the same min() and max() values
for each year. 

```{r Get legend of the arranged plot}

leg <- get_legend(SPEI_2018_plot)

# Convert to a ggplot object
leg <- as_ggplot(leg)
leg

png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240422_SHAP_predictor_maps/20240423_SPEI_sameyear_SHAP_plots/leg_SPEI_SHAP_allyears.png", 
    width = 6.1, height = 6, 
    units = "cm", res = 300)
leg
dev.off()


```
