---
title: 'Fig 5: Mean SHAP values of five most important predictors'
author: "Katharina Horn"
date: "2024-05-13"
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

library(ggspatial)
#library(geobuffer)
library(rgeos)

library(cowplot) # get_legend() function
library(ggpubr) # as_ggplot() function

library(tools)
library(tibble)

```

# Import Brandenburg and Berlin border

Keeping Berlin in the plots.

``` {r Load Berlin-Brandenburg border}

bb <- terra::vect("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp") # for cropping of data frames

# Load as sf object
bb_st <- st_read("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp") # for mapping

```

## Load five predictor data frames


```{r load data frames of mean SHAP values of five most important predictors}

broadleaf_path = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/broadleaf_perc_mean_SHAP_df.csv"
canopy_path = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/canopyheight_mean_SHAP_df.csv"
ET_path = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/ET_1year_mean_SHAP_df.csv"
SPEI_path = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/SPEI_sameyear_mean_SHAP_df.csv"
TCD_path = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/TCD_mean_SHAP_df.csv"

# first two columns are x and y ´
# 8th column is mean_shap across all years

broadleaf_df <- data.table::fread(broadleaf_path, select = c(1,2,8), header = TRUE)
canopy_df <- data.table::fread(canopy_path, select = c(1,2,8), header = TRUE)
ET_df <- data.table::fread(ET_path, select = c(1,2,8), header = TRUE)
SPEI_df <- data.table::fread(SPEI_path, select = c(1,2,8), header = TRUE)
TCD_df <- data.table::fread(TCD_path, select = c(1,2,8), header = TRUE)

glimpse(broadleaf_df)
# looks fine


```

## Raster conversion is not needed, since I am plotting with ggplot() and therefore need data frames instead of raster files. 

```{r convert to raster }

# broadleaf_SHAP_rast = terra::rast(broadleaf_df, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
# canopy_SHAP_rast = terra::rast(canopy_df, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
# ET_SHAP_rast = terra::rast(ET_df, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
# SPEI_SHAP_rast = terra::rast(SPEI_df, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
# TCD_SHAP_rast = terra::rast(TCD_df, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")
# 
# # the reason I converted to lonlat is to include the scalebar (it didnt work with the other projection)
# broadleaf_SHAP_rast_lonlat = terra::project(x = broadleaf_SHAP_rast, y = "+proj=longlat +datum=WGS84 +no_defs")
# canopy_SHAP_rast_lonlat = terra::project(x = canopy_SHAP_rast, y = "+proj=longlat +datum=WGS84 +no_defs")
# ET_SHAP_rast_lonlat = terra::project(x = ET_SHAP_rast, y = "+proj=longlat +datum=WGS84 +no_defs")
# SPEI_SHAP_rast_lonlat = terra::project(x = SPEI_SHAP_rast, y = "+proj=longlat +datum=WGS84 +no_defs")
# TCD_SHAP_rast_lonlat = terra::project(x = TCD_SHAP_rast, y = "+proj=longlat +datum=WGS84 +no_defs")



# # convert to raster::raster for plotting
# SHAP_r = raster::raster(SHAP_rast_lonlat)
# 
# raster::plot(SHAP_r, colNA = "lightgray", col=colorRampPalette(c("#460000", "#930c0c", "#ff2100", "#ff980c",
#                                  "#fcf599", "#00ff9b", "#39b5ff","#002ae7",
#                                 "#020d47"))(255),
#              legend.args = list(text = "SHAP value", line = 0.5, cex = 1, adj = 0.25))
# 
# minValue(SHAP_r) # -0.6971202
# maxValue(SHAP_r) # 1.486262


```

# Define color scale

```{r Define color scale}

colors <- colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255)

```

# Plot mean SHAP of all five predictors

```{r Plot mean SHAP of all five predictors separately}

# Plot TCD_SHAP
color_limits_TCD <- c(-1, 1)
# used for tcd and broadleaf_perc


# Plot SPEI_SHAP
color_limits_SPEI <- c(-0.3, 0.3)
# used for SPEI, ET, and canopyheight


TCD_SHAP <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = mean_SHAP), data = TCD_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_TCD, guide = "colorbar", breaks = c(color_limits_TCD[1], 0, color_limits_TCD[2]), oob = scales::squish) +    # set legend colours
  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal", legend.position = "bottom") +
  theme_void() +
  theme(legend.position = "bottom")                                                           # no theme

# Plot broadleaf_SHAP
broadleaf_SHAP <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = mean_SHAP), data = broadleaf_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_TCD, guide = "colorbar", breaks = c(color_limits_TCD[1], 0, color_limits_TCD[2]), oob = scales::squish) +    labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal") +
  theme_void() +
  theme(legend.position = "bottom")                                                           # no theme


# Plot canopy_SHAP
canopy_SHAP <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = mean_SHAP), data = canopy_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]), oob = scales::squish) +                            # set legend colours  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal") +
  theme_void() +
  theme(legend.position = "bottom")                                                           # no theme

# Plot SPEI_SHAP
SPEI_SHAP <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = mean_SHAP), data = SPEI_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]), oob = scales::squish) +                            # set legend colours  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme(legend.direction = "horizontal", legend.box = "horizontal") +
  theme_void()    +
  theme(legend.position = "bottom")                                                        # no theme

# Plot ET_SHAP
ET_SHAP <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = mean_SHAP), data = ET_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors, limits = color_limits_SPEI, guide = "colorbar", breaks = c(color_limits_SPEI[1], 0, color_limits_SPEI[2]), oob = scales::squish) +                            # set legend colours
  labs(x = NULL, y = NULL, fill = "SHAP values")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                      # add scale bar
  theme_void() +                                                           # no theme
  theme(legend.direction = "horizontal", legend.box = "horizontal") +
  theme(legend.position = "bottom")

```

# Get legend of the plots

I can extract a legend from one of the plots since all of them are based on the same min() and max() values
for each year. 

```{r Get legend of one of the plots to use as joint legend for all plots}

leg1 <- get_legend(TCD_SHAP)
leg2 <- get_legend(SPEI_SHAP)

# Convert to a ggplot object
leg1 <- as_ggplot(leg1)
leg2 <- as_ggplot(leg2)

```

# Arrange and export plots

In Inkscape, I will add the label headings (aligning did not work out well within ggarrange() and I will add the scales as well. 

```{r Arrange resulting plots}

# Create the layout
png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240425_mostimp_SHAP_avg_maps/predictors_avgSHAP_map_new.png", width = 17.5, height = 12, units = "cm", res = 300)
predictors_avgSHAP <- ggarrange(
  ggarrange(TCD_SHAP, broadleaf_SHAP, ncol = 3, nrow = 1, legend = "none"),
  ggarrange(canopy_SHAP, SPEI_SHAP, ET_SHAP, ncol = 3, nrow = 1, legend = "none"),
  ncol = 1, nrow = 2
)

predictors_avgSHAP

dev.off()


```
# Export legends

Lastly, I am saving the legends as png files so that I can combine everything in Inkscape. 

```{r Save legends as png files}

# Export legends as png files for arranging in inkscape
png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240425_mostimp_SHAP_avg_maps/leg1_TCD.png", width = 5.9, height = 2, units = "cm", res = 300)
leg1
dev.off()

# Export legends as png files for arranging in inkscape
png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240425_mostimp_SHAP_avg_maps/leg2_SPEI.png", width = 5.9, height = 2, units = "cm", res = 300)
leg2
dev.off()

# I am 
```
