---
title: "Dynamic predictors"
author: "Katharina Horn"
date: "2024-05-22"
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
#library(raster)
#library(rgdal)
library(lubridate)
#library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
#library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

library(ggspatial)
library(cowplot)
library(ggpubr)
#library(geobuffer)

library(tools)
library(tibble)

```


# Load predictor rasters

TOTAL number of predictors: 41 (31 unique predictors - if we don’t count “time lags”)
Therefore, I am loading the rasters of the dynamic predictors.
Those are SPEI, SMI, SSM, SWI, LST, Air temperature, Evapotranspiration, direct radiation, ground water level, and ground water level trend
for the years of 2017 to 2022 (six layers per predictor). This makes a total of 60 plots. 

```{r load annual predictor files}

# Meteorology and Drought indices
spei <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/SPEI_annual_2017_to_2022.tif")
smi <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/SMI_annual_2017_to_2022.tif")
ssm <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/SSM_annual_2017_to_2022.tif")
swi <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/SWI_15cm_annual_2017_to_2022.tif")
lst <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/LST_annual_2017_to_2022.tif")
airtemp <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/airtemp_annual_2017_to_2022.tif")
et <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/ET_annual_2017_to_2022.tif")
rad <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/rad_direct_annual_2017_to_2022.tif")
gw_level <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/groundwater_annual_2017_to_2022.tif")
gw_trend <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/gw_trend_annual_2017_to_2022.tif")

```


```{r load jan-jul predictor files}

airtemp_jj <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Jan_to_July/airtemp_JanJuly_2017_to_2022.tif")
ET_jj <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Jan_to_July/ET_JanJuly_2017_to_2022.tif")
rad_jj <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Jan_to_July/rad_direct_JanJuly_2017_to_2022.tif")
SMI_jj <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Jan_to_July/SMI_JanJuly_2017_to_2022.tif")
SPEI_jj <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Jan_to_July/SPEI_JanJuly_2017_to_2022.tif")
SSM_jj <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Jan_to_July/SSM_JanJuly_2017_to_2022.tif")
SWI_jj <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Jan_to_July/SWI_15cm_JanJuly_2017_to_2022.tif")


```


```{r Define color scale}

colors <- colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255)

colors_rev <- rev(colors)

```


```{r Plot Page 1 of dynamic predictors}

# spei
spei_layers <- c(spei[[1]], spei[[2]], spei[[3]], spei[[4]], spei[[5]], spei[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Annual/S1_dyn_pred_spei2.png", 
    height = 3.5, width = 20, units = "cm", res = 300)

terra::panel(spei_layers, nc = 6, nr = 1, range = c(-1.6, 1.6), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()

# smi
smi_layers <- c(smi[[1]], smi[[2]], smi[[3]], smi[[4]], smi[[5]], smi[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Annual/S1_dyn_pred_smi.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(smi_layers, nc = 6, nr = 1, range = c(0, 0.8), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()

# ssm
ssm_layers <- c(ssm[[1]], ssm[[2]], ssm[[3]], ssm[[4]], ssm[[5]], ssm[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Annual/S1_dyn_pred_ssm.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(ssm_layers, nc = 6, nr = 1, range = c(0, 100), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()

# swi
swi_layers <- c(swi[[1]], swi[[2]], swi[[3]], swi[[4]], swi[[5]], swi[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Annual/S1_dyn_pred_swi.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(swi_layers, nc = 6, nr = 1, range = c(25, 86), 
             main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()

# lst
lst_layers <- c(lst[[1]], lst[[2]], lst[[3]], lst[[4]], lst[[5]], lst[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Annual/S1_dyn_pred_lst.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(lst_layers, nc = 6, nr = 1, range = c(10, 50), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()


```

```{r Plot Page 2 of dynamic predictors}

# airtemp
airtemp_layers <- c(airtemp[[1]], airtemp[[2]], airtemp[[3]], airtemp[[4]], airtemp[[5]], airtemp[[6]])

airtemp_layers
png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/S1_dyn_pred_airtemp.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(airtemp_layers, nc = 6, nr = 1, range = c(8.9, 11.9), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()

# et
et_layers <- c(et[[1]], et[[2]], et[[3]], et[[4]], et[[5]], et[[6]])

et_layers
png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/S1_dyn_pred_et.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(et_layers, nc = 6, nr = 1, range = c(323, 510), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()

# rad
rad_layers <- c(rad[[1]], rad[[2]], rad[[3]], rad[[4]], rad[[5]], rad[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/S1_dyn_pred_rad.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(rad_layers, nc = 6, nr = 1, range = c(440, 732), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()

# gw_level
gw_level_layers <- c(gw_level[[1]], gw_level[[2]], gw_level[[3]], gw_level[[4]], gw_level[[5]], gw_level[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/S1_dyn_pred_gw_level.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(gw_level_layers, nc = 6, nr = 1, range = c(0.7, 18.3), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()

# gw_trend
gw_trend_layers <- c(gw_trend[[1]], gw_trend[[2]], gw_trend[[3]], gw_trend[[4]], gw_trend[[5]], gw_trend[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/S1_dyn_pred_gw_trend.png", 
    height = 3.5, width = 20, units = "cm", res = 300)
terra::panel(gw_trend_layers, nc = 6, nr = 1, range = c(-0.5, 0.35), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()


```



```{r Plot Page 3: Jan_July predictors}

# SPEI_jj
SPEI_jj_layers <- c(SPEI_jj[[1]], SPEI_jj[[2]], SPEI_jj[[3]], SPEI_jj[[4]], SPEI_jj[[5]], SPEI_jj[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Jan_July/S1_dyn_pred_SPEI_jj.png", 
    height = 4.1, width = 20, units = "cm", res = 300)

terra::panel(SPEI_jj_layers, nc = 6, nr = 1, range = c(-1.5, 1.15), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()


# SSM_jj
SSM_jj_layers <- c(SSM_jj[[1]], SSM_jj[[2]], SSM_jj[[3]], SSM_jj[[4]], SSM_jj[[5]], SSM_jj[[6]])

png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Jan_July/S1_dyn_pred_SSM_jj.png", 
    height = 4.1, width = 20, units = "cm", res = 300)

terra::panel(SSM_jj_layers, nc = 6, nr = 1, range = c(0, 100), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()


# SMI_jj
SMI_jj_layers <- c(SMI_jj[[1]], SMI_jj[[2]], SMI_jj[[3]], SMI_jj[[4]], SMI_jj[[5]], SMI_jj[[6]])
SMI_jj_layers
png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Jan_July/S1_dyn_pred_SMI_jj.png", 
    height = 4.1, width = 20, units = "cm", res = 300)

terra::panel(SMI_jj_layers, nc = 6, nr = 1, range = c(0, 1), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()

# SWI_jj
SWI_jj_layers <- c(SWI_jj[[1]], SWI_jj[[2]], SWI_jj[[3]], SWI_jj[[4]], SWI_jj[[5]], SWI_jj[[6]])
SWI_jj_layers
png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Jan_July/S1_dyn_pred_SWI_jj.png", 
    height = 4.1, width = 20, units = "cm", res = 300)

terra::panel(SWI_jj_layers, nc = 6, nr = 1, range = c(20, 87), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors_rev, axes = FALSE)
dev.off()

```

```{r plot Page 4: Jan_july dynamic predictors}

# airtemp_jj
airtemp_jj_layers <- c(airtemp_jj[[1]], airtemp_jj[[2]], airtemp_jj[[3]], airtemp_jj[[4]], airtemp_jj[[5]], airtemp_jj[[6]])
airtemp_jj_layers
png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Jan_July/S1_dyn_pred_airtemp_jj.png", 
    height = 5.5, width = 20, units = "cm", res = 300)

terra::panel(airtemp_jj_layers, nc = 6, nr = 1, range = c(8, 12), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()

# ET_jj
ET_jj_layers <- c(ET_jj[[1]], ET_jj[[2]], ET_jj[[3]], ET_jj[[4]], ET_jj[[5]], ET_jj[[6]])
ET_jj_layers
png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Jan_July/S1_dyn_pred_ET_jj.png", 
    height = 5.5, width = 20, units = "cm", res = 300)

terra::panel(ET_jj_layers, nc = 6, nr = 1, range = c(230, 340), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()

# rad_jj
rad_jj_layers <- c(rad_jj[[1]], rad_jj[[2]], rad_jj[[3]], rad_jj[[4]], rad_jj[[5]], rad_jj[[6]])
rad_jj_layers
png(filename = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240515_supplement_figure_predictors/Dynamic/Jan_July/S1_dyn_pred_rad_jj.png", 
    height = 5.5, width = 20, units = "cm", res = 300)

terra::panel(rad_jj_layers, nc = 6, nr = 1, range = c(320, 525), main = c("2017", "2018", "2019", "2020", "2021", "2022"), 
             loc.main = "bottomleft", col = colors, axes = FALSE)
dev.off()


```


