---
title: "NDMI_anomaly mapping"
author: "Katharina Horn"
date: '2024-05-10'
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(ggpubr)  # plotting multiple maps together
library(patchwork) # to use plot_spacer()
library(cowplot) # to use plot_grid()

library(terra)

library(ggspatial)
#library(geobuffer)
library(rgeos)

library(tools)
library(tibble)

library(scales)

```


```{r Load NDMI anomaly rasters}

NDMI_2018 <- terra::rast("D:/Katharina/CliWaC/03_PROCESSED DATA/Vegetation/NDMI/Anomaly Rasters with TCD/NDMI_2018_anom_forest.tif")
NDMI_2019 <- terra::rast("D:/Katharina/CliWaC/03_PROCESSED DATA/Vegetation/NDMI/Anomaly Rasters with TCD/NDMI_2019_anom_forest.tif")
NDMI_2020 <- terra::rast("D:/Katharina/CliWaC/03_PROCESSED DATA/Vegetation/NDMI/Anomaly Rasters with TCD/NDMI_2020_anom_forest.tif")
NDMI_2021 <- terra::rast("D:/Katharina/CliWaC/03_PROCESSED DATA/Vegetation/NDMI/Anomaly Rasters with TCD/NDMI_2021_anom_forest.tif")
NDMI_2022 <- terra::rast("D:/Katharina/CliWaC/03_PROCESSED DATA/Vegetation/NDMI/Anomaly Rasters with TCD/NDMI_2022_anom_forest.tif")

#NDMI_stack <- raster::stack(NDMI_2018, NDMI_2019, NDMI_2020, NDMI_2021, NDMI_2022)
#NDMI_stack <- c(NDMI_2018, NDMI_2019, NDMI_2020, NDMI_2021, NDMI_2022)
#names(NDMI_stack) <- c("NDMI_2018", "NDMI_2019", "NDMI_2020", "NDMI_2021", "NDMI_2022")
#NDMI_stack 

```

# Reproject Berlin-Brandenburg border

No need to redo this, since I have already run this part of code once. 

```{r Reproject Berlin-Brandenburg border}

bb <- terra::vect("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border.shp")

bb_proj <- terra::project(bb, NDMI_2018)
bb_proj
plot(bb_proj)

writeVector(bb_proj, "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp")

```

# Load Berlin-Brandenburg border

``` {r Load Berlin-Brandenburg border}

bb <- terra::vect("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp") # for cropping of data frames

# Load as sf object
bb_st <- st_read("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp") # for mapping

```

```{r Load DEM layer to use for excluding Berlin NDMI values}

dem <- terra::rast("D:/Katharina/CliWaC/03_PROCESSED DATA/Static_data_proj/dem_proj.tif")

dem_proj <- terra::project(dem, NDMI_2018_crop)

```


```{r Mask NDMI layers to dem to exclude values within Berlin}

NDMI_2018_mask <- terra::crop(NDMI_2018, dem, mask = TRUE)
NDMI_2019_mask <- terra::crop(NDMI_2019, dem, mask = TRUE)
NDMI_2020_mask <- terra::crop(NDMI_2020, dem, mask = TRUE)
NDMI_2021_mask <- terra::crop(NDMI_2021, dem, mask = TRUE)
NDMI_2022_mask <- terra::crop(NDMI_2022, dem, mask = TRUE)


```


```{r crop NDMI layers to BB border}

NDMI_2018_crop <- raster::crop(NDMI_2018_mask, bb)
NDMI_2019_crop <- raster::crop(NDMI_2019_mask, bb)
NDMI_2020_crop <- raster::crop(NDMI_2020_mask, bb)
NDMI_2021_crop <- raster::crop(NDMI_2021_mask, bb)
NDMI_2022_crop <- raster::crop(NDMI_2022_mask, bb)

```


```{r convert rasters into dataframes}

# Convert rasters into data frames
NDMI_2018_df <- as.data.frame(NDMI_2018_crop, xy = TRUE)
NDMI_2019_df <- as.data.frame(NDMI_2019_crop, xy = TRUE)
NDMI_2020_df <- as.data.frame(NDMI_2020_crop, xy = TRUE)
NDMI_2021_df <- as.data.frame(NDMI_2021_crop, xy = TRUE)
NDMI_2022_df <- as.data.frame(NDMI_2022_crop, xy = TRUE)

```


```{r Create subsets of data frames to exclude outliers}

NDMI_2018_df_filtered <- subset(NDMI_2018_df, AVG >= -30 & AVG <= 30)
NDMI_2019_df_filtered <- subset(NDMI_2019_df, AVG >= -30 & AVG <= 30)
NDMI_2020_df_filtered <- subset(NDMI_2020_df, AVG >= -30 & AVG <= 30)
NDMI_2021_df_filtered <- subset(NDMI_2021_df, AVG >= -30 & AVG <= 30)
NDMI_2022_df_filtered <- subset(NDMI_2022_df, AVG >= -30 & AVG <= 30)

```


```{r Define color scale}

colors <- colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255)
colors_rev <- c("#7F0000", "#F82602", "#F36802", "#ff980c", "#fcf599", "#00ff9b", "#39b5ff", "#002ae7", "#020d47")

```

# Plot NDMI for each year

I decided not to plot the north arrow, because it does not look good on the final map. I will add it with Inkscape afterwards. 

```{r Plot NDMI for all years}

# Plot NDMI_anomaly 2018
map_ndmi_2018 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = AVG), data = NDMI_2018_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors_rev, limits = c(-30, 30), oob = scales::squish) +      # set legend colours; oob = scales::squish is setting all values out of the limits to the nearest colour respectively
  labs(x = NULL, y = NULL, fill = "NDMI anomaly (%)")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
#  annotation_north_arrow(location = "tr", which_north = "grid", height = unit(0.5, "cm"), width = unit(0.35, "cm")) +          # add north arrow
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2019
map_ndmi_2019 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = AVG), data = NDMI_2019_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors_rev, limits=c(-30,30), oob = scales::squish) +                                  # set legend colours
  labs(x = NULL, y = NULL, fill = "NDMI anomaly 2019 (%)")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2020
map_ndmi_2020 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = AVG), data = NDMI_2020_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors_rev, limits=c(-30,30), oob = scales::squish) +                                  # set legend colours
  labs(x = NULL, y = NULL, fill = "NDMI anomaly 2020 (%)")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2021
map_ndmi_2021 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = AVG), data = NDMI_2021_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors_rev, limits=c(-30,30), oob = scales::squish) +                                  # set legend colours
  labs(x = NULL, y = NULL, fill = "NDMI anomaly 2021 (%)")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2022
map_ndmi_2022 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = AVG), data = NDMI_2022_df) +       # add filtered raster data set (excl. outliers)
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_gradientn(colours = colors_rev, limits=c(-30,30), oob = scales::squish) +                                  # set legend colours
  labs(x = NULL, y = NULL, fill = "NDMI anomaly 2022 (%)")  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

```

# Check results before plotting everything.

```{r Check resulting plot for 2018}

map_ndmi_2018

# png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240510_NDMI_anomaly/NDMI_2018.png", width = 17.5, height = 12, units = "cm", res = 300)
# map_ndmi_2018
# dev.off()

```

# Get legend of one of the plots

I can extract a legend from one of the plots since all of them are based on the same min() and max() values
for each year. 

```{r Get legend of one of the plots to use as joint legend for all plots}

leg <- get_legend(map_ndmi_2018)

# Convert to a ggplot object
leg <- as_ggplot(leg)
leg

```


```{r Arrange resulting plots}

png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240510_NDMI_anomaly/NDMI_anomaly_figure_scale30.png", width = 17.5, height = 12, units = "cm", res = 300)

ndmi <- ggarrange(map_ndmi_2018, map_ndmi_2019, map_ndmi_2020, map_ndmi_2021, map_ndmi_2022, leg,
          labels = c("2018", "2019", "2020", "2021", "2022", " "), 
          ncol = 3, nrow = 2, legend = "none")
ndmi

dev.off()

```


## Alternative plotting option based on raster package. (To be continued if needed.)

```{r import Berlin-Brandenburg border }
# 
# BB_border = readOGR(dsn = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border",
#                         layer = "Brandenburg_border")
# 
# crs(BB_border)
# # CRS arguments: +proj=longlat +datum=WGS84 +no_defs 
# 
# plot(BB_border)
# 
# crs <- crs(NDMI_2018)
# 
# # +proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs
# BB_LAEA = sp::spTransform(BB_border, CRSobj = crs)
# BB_border = BB_LAEA


```

```{r plot NDMI_anomaly}

# png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240510_NDMI_anomaly/NDMI_anomaly_figure.png", width = 17.5, height = 12, units = "cm", res = 300)
# 
# # Set outer margins to accommodate legend
# par(oma = c(0, 0, 0, 3)) # I added this, because otherwise the legend was cut off. 
# 
# raster::plot(NDMI_stack_crop, 
#              col = colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255),
#              zlim=c(-7,5),
#              legend.args = list(text = "NDMI anomaly", line = 0.5, cex = 1, adj = 0.25),
#              legend = TRUE, axes = FALSE, box = FALSE)
# 
# #plot(mostimp_lonlat, col = c("green3", "magenta3", "gold2", "dodgerblue", "forestgreen"),
# #     legend.only = TRUE, horizontal = FALSE , smallplot = c(0.87, 0.9, 0.17, 0.86),
# #     axis.args = list(cex.axis = 1.4), 
# #	legend.args = list(text = "Most contributing predictor", line = 0.5, cex = 1, adj = 0.25))
# 
# # add scalebar 
# raster::scalebar(d = 50, # distance in km,
#                  xy = c(11.5, 51.5),
#                  type = "bar",
#                  divs = 2,
#                  below = "km",
#                  lonlat = TRUE,
#                  adj = c(0.5, -1.15),
#                  cex = 1.2)
# 
# raster::plot(BB_border, add = TRUE)
# 
# 
# dev.off()
# 
# 


```