---
title: "Average SHAP values (LST_sameyear)"
author: "Stenka Vulova"
date: '2024-04-23'
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

library(ggspatial)
#library(geobuffer)
library(rgeos)

library(tools)
library(tibble)

```

```{r import SHAP df names }

# column names of SHAP dfs
cols_SHAPdf = read.table("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/colnames_SHAPdf.csv", sep = ",", header = FALSE)

```

# Select a predictor

Select which predictor you want to analyze (and get what column # it is in SHAP dfs)

```{r select the top predictors }

pred_name = "LST_sameyear"

# Match the names in pred_name with the values in the "V1" column of cols_SHAPdf
matched_rows <- match(pred_name, cols_SHAPdf$V1, nomatch = 0)

# Filter out the matched row numbers
matched_rows <- matched_rows[matched_rows > 0]

# Print the list of row numbers
print(matched_rows) # 32

# 32: LST_sameyear

# looks good. 

```

# Load the SHAP df 2018 

```{r import the 2018 df }

tic("load 2018 SHAP df")

shap_df_path_2018 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2018_SHAP.csv"

# read only a selected column name e.g. predictor
shap_df_2018 <- data.table::fread(shap_df_path_2018, select = matched_rows, header = TRUE)

toc()

tic.clearlog()
# load 2018 SHAP df: 435.52 sec elapsed

glimpse(shap_df_2018)
# Rows: 10,406,171
#Columns: 1

# assign the column name as the selected predictor 
colnames(shap_df_2018) = "2018"

```

# Coordinates df 2018 

I am loading the coordinates. 

```{r load coords 2018 }

# reference to see column names in true_class dfs
#true_class <- mod_df[, (names(mod_df) %in% c("x", "y", "coord_code", "true_class", "NDMI_anomaly"))]

true_class_path_2018 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230927_modeling_df/true_class/true_class_2018.csv"

# first two columns are x and y 

coords_df_2018 <- data.table::fread(true_class_path_2018, select = c(1,2), header = TRUE)

glimpse(coords_df_2018)

# Rows: 10,406,171
# Columns: 2
# x
# y

```

# Make into one df 2018

```{r new df with coords 2018 }

SHAPdf_coords_2018 = cbind(coords_df_2018, shap_df_2018)

glimpse(SHAPdf_coords_2018)

# Rows: 10,406,171
#Columns: 3
# x
# y
# '2018'

```

# Load the SHAP df 2019

```{r import the 2019 df }

tic("load 2019 SHAP df")

shap_df_path_2019 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2019_SHAP.csv"

# read only a selected column name e.g. predictor
shap_df_2019 <- data.table::fread(shap_df_path_2019, select = matched_rows, header = TRUE)

toc()

tic.clearlog()
# load 2019 SHAP df: 246.26 sec elapsed

glimpse(shap_df_2019)
# Rows: 10,437,444
#Columns: 1

colnames(shap_df_2019) = "2019"

```

# Coordinates df 2019 

I am loading the coordinates. 

```{r load coords 2019 }

# reference to see column names in true_class dfs
#true_class <- mod_df[, (names(mod_df) %in% c("x", "y", "coord_code", "true_class", "NDMI_anomaly"))]

true_class_path_2019 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230927_modeling_df/true_class/true_class_2019_KH.csv"

# first two columns are x and y 

coords_df_2019 <- data.table::fread(true_class_path_2019, select = c(1,2), header = TRUE)

glimpse(coords_df_2019)

# Rows: 10,437,444
# Columns: 2
# x
# y

```

# Make into one df 2019

```{r new df with coords 2019 }

SHAPdf_coords_2019 = cbind(coords_df_2019, shap_df_2019)

glimpse(SHAPdf_coords_2019)

# Rows: 10,437,444
#Columns: 3
# x
# y
# LST_sameyear_2019

```

# Load the SHAP df 2020 

```{r import the 2020 df }

tic("load 2020 SHAP df")

shap_df_path_2020 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2020_SHAP.csv"

# read only a selected column name e.g. predictor
shap_df_2020 <- data.table::fread(shap_df_path_2020, select = matched_rows, header = TRUE)

toc()

tic.clearlog()
# load 2020 SHAP df: 314.75 sec elapsed

glimpse(shap_df_2020)
# Rows: 10,434,985
#Columns: 1

# assign the column name as the selected predictor 
colnames(shap_df_2020) = "2020"

```

# Coordinates df 2020 

I am loading the coordinates. 

```{r load coords 2020 }

# reference to see column names in true_class dfs
#true_class <- mod_df[, (names(mod_df) %in% c("x", "y", "coord_code", "true_class", "NDMI_anomaly"))]

true_class_path_2020 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230927_modeling_df/true_class/true_class_2020.csv"

# first two columns are x and y 

coords_df_2020 <- data.table::fread(true_class_path_2020, select = c(1,2), header = TRUE)

glimpse(coords_df_2020)

# Rows: 10,434,985
# Columns: 2
# x
# y

```

# Make into one df 2020

```{r new df with coords 2020 }

SHAPdf_coords_2020 = cbind(coords_df_2020, shap_df_2020)

glimpse(SHAPdf_coords_2020)

# Rows: 10,434,985
# Columns: 3
# x
# y
# LST_sameyear_2020

```

# Load the SHAP df 2021 

```{r import the 2021 df }

tic("load 2021 SHAP df")

shap_df_path_2021 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2021_SHAP.csv"

# read only a selected column name e.g. predictor
shap_df_2021 <- data.table::fread(shap_df_path_2021, select = matched_rows, header = TRUE)

toc()

tic.clearlog()
# load 2021 SHAP df: 417.48 sec elapsed

glimpse(shap_df_2021)
# Rows: 9,624,727
#Columns: 1

colnames(shap_df_2021) = "2021"


```

# Coordinates df 2021 

I am loading the coordinates. 

```{r load coords 2021 }

# reference to see column names in true_class dfs
#true_class <- mod_df[, (names(mod_df) %in% c("x", "y", "coord_code", "true_class", "NDMI_anomaly"))]

true_class_path_2021 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230927_modeling_df/true_class/true_class_2021.csv"

# first two columns are x and y 

coords_df_2021 <- data.table::fread(true_class_path_2021, select = c(1,2), header = TRUE)

glimpse(coords_df_2021)

# Rows: 9,624,727
# Columns: 2
# x
# y

```

# Make into one df 2021

```{r new df with coords 2021 }

SHAPdf_coords_2021 = cbind(coords_df_2021, shap_df_2021)

glimpse(SHAPdf_coords_2021)

# Rows: 9,624,727
#Columns: 3
# x
# y
# LST_sameyear_2021

```

# Load the SHAP df 2022 

```{r import the 2022 df }

tic("load 2022 SHAP df")

shap_df_path_2022 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/SHAP.df_NEW/dfs/X_all_2022_SHAP.csv"

# read only a selected column name e.g. predictor
shap_df_2022 <- data.table::fread(shap_df_path_2022, select = matched_rows, header = TRUE)

toc()

tic.clearlog()
# load 2022 SHAP df: 226.43 sec elapsed

glimpse(shap_df_2022)
# Rows: 10,432,546
#Columns: 1

colnames(shap_df_2022) = "2022"

```

# Coordinates df 2022 

I am loading the coordinates. 

```{r load coords 2022 }

# reference to see column names in true_class dfs
#true_class <- mod_df[, (names(mod_df) %in% c("x", "y", "coord_code", "true_class", "NDMI_anomaly"))]

true_class_path_2022 = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230927_modeling_df/true_class/true_class_2022.csv"

# first two columns are x and y 

coords_df_2022 <- data.table::fread(true_class_path_2022, select = c(1,2), header = TRUE)

glimpse(coords_df_2022)

# Rows:  10,432,546
# Columns: 2
# x
# y

```

# Make into one df 2022

```{r new df with coords 2022 }

SHAPdf_coords_2022 = cbind(coords_df_2022, shap_df_2022)

glimpse(SHAPdf_coords_2022)

# Rows: 10,406,171
#Columns: 3
# x
# y
# LST_sameyear_2022

```

# Join all

```{r join all }

list_dfs = list(SHAPdf_coords_2018, SHAPdf_coords_2019, SHAPdf_coords_2020, SHAPdf_coords_2021, SHAPdf_coords_2022)

tic("merge dfs")
merged_SHAPdf = plyr::join_all(dfs = list_dfs, type = "full")

toc()

tictoc::tic.clearlog()

# merge dfs: 163.83 sec elapsed

glimpse(merged_SHAPdf)

# Rows: 10,439,063
# Columns: 7

summary(merged_SHAPdf)

```



# Mean of SHAP values

```{r mean shap }

merged_SHAPdf$mean_SHAP = rowMeans(subset(merged_SHAPdf, select = c("2018", "2019", "2020", "2021", "2022")),
                                          na.rm = TRUE)


# save this df 
write.csv(merged_SHAPdf, "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/LST_sameyear_mean_SHAP_df.csv", row.names = FALSE)

glimpse(merged_SHAPdf)
# Rows: 10,439,000
# Columns: 8

colnames(merged_SHAPdf)
# [1] "x"         "y"         "2018"      "2019"      "2020"      "2021"      "2022"      "mean_SHAP"

TEST = read.csv("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/LST_sameyear_mean_SHAP_df.csv")

# Rows: 10,439,063
# Columns: 8

```

# Convert to raster TEST

```{r convert to raster }

shap_df = merged_SHAPdf[, c("x", "y", "mean_SHAP")]

SHAP_rast = terra::rast(shap_df, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

# the reason I converted to lonlat is to include the scalebar (it didnt work with the other projection)
SHAP_rast_lonlat = terra::project(x = SHAP_rast, y = "+proj=longlat +datum=WGS84 +no_defs")

# convert to raster::raster for plotting
SHAP_r = raster::raster(SHAP_rast_lonlat)

raster::plot(SHAP_r, colNA = "lightgray", col = colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255),
             legend.args = list(text = "SHAP value", line = 0.5, cex = 1, adj = 0.25))

minValue(SHAP_r) # -0.3587276
maxValue(SHAP_r) # 0.6546246

# it works, cool 

```
