---
title: "Resample TWI to MODIS"
author: "Stenka Vulova"
date: '2023-06-15'
output: html_document
---

New TWI raster from Katharina 

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

library(ggspatial)
#library(geobuffer)
library(rgeos)

library(tools)

```

# Import datasets

## TWI

I will resample TWI to MODIS NDVI (1 km resolution).
This version has not been reclassified (it has 13 classes, 1-13).

```{r import TWI }

TWI = terra::rast("D:/Stenka_Cliwac/Topic_1/04_PROCESSED_DATA/20230530_Topography/TWI/TWI_merged_wgs84.tif")
terra::crs(TWI)

terra::plot(TWI, col = colorRampPalette(c("#460000", "#930c0c", "#ff2100", "#ff980c",
                                 "#fcf599", "#00ff9b", "#39b5ff","#002ae7",
                                "#020d47"))(255))

# , range = c (0,360)

TWI
#class       : SpatRaster 
#dimensions  : 22154, 47464, 1  (nrow, ncol, nlyr)
#resolution  : 0.0001273666, 0.0001273666  (x, y)
#extent      : 9.682033, 15.72736, 50.88576, 53.70744  (xmin, xmax, ymin, ymax)
#coord. ref. : lon/lat WGS 84 (EPSG:4326) 
#source      : TWI_merged_wgs84.tif 
#name        : TWI_merged_wgs84 
#min value   :        -8.478123 
#max value   :         38.07537

```

## MODIS NDVI

```{r import MODIS NDVI }

NDVI = terra::rast("D:/Stenka_Cliwac/Topic_1/04_PROCESSED_DATA/20230327_MODIS_NDVI_anom_forest/NDVI_anomaly_2019_forest.tif")
terra::crs(NDVI)
terra::crs(TWI)

terra::plot(NDVI)

NDVI

```


## Import Brandenburg and Berlin border

```{r import border }

BB_border = readOGR(dsn = "D:/Stenka_Cliwac/Topic_1/04_PROCESSED_DATA/20230214_Brandenburg_border",
                        layer = "BB_border_merged")

crs(BB_border)
# CRS arguments: +proj=longlat +datum=WGS84 +no_defs 

plot(BB_border)

BB_vect = terra::vect(BB_border)

terra::plot(BB_vect)

```

# Function project_crop_save

1) Project a given raster to the NDVI raster (MODIS)
2) Crop to the Brandenburg border  
3) return or save the raster

```{r function project_crop_save }

# x: Spatraster 
# y: SpatRaster to project to; e.g. MODIS NDVI in this example 
# border: Spatvector type of polygon to mask the raster to; the Brandenburg border in this example 
# filepath: ENTIRE filepath to save tif file to, e.g. "D:/Stenka_Cliwac/Topic_1/04_PROCESSED_DATA/20230522_static_data_proj/TWI_proj.tif"

project_crop_save = function(x, y, border, filepath) {
  
  x_proj = terra::project(x = x, y = y) # project
  
  mask = terra::mask(x_proj, border) 
  
  terra::writeRaster(x = mask, filename = filepath )
}

```

# Run project_crop_save

```{r run function }

project_crop_save(x = TWI, y = NDVI, border = BB_vect,
                  filepath = "D:/Stenka_Cliwac/Topic_1/04_PROCESSED_DATA/20230522_static_data_proj/TWI_proj.tif")

```

# Check result

```{r check result }

TWI_proj = terra::rast("D:/Stenka_Cliwac/Topic_1/04_PROCESSED_DATA/20230522_static_data_proj/TWI_proj.tif")
terra::plot(TWI_proj, col = colorRampPalette(c("#460000", "#930c0c", "#ff2100", "#ff980c",
                                 "#fcf599", "#00ff9b", "#39b5ff","#002ae7",
                                "#020d47"))(255))

stack = c(TWI_proj, NDVI)

terra::plot(stack)

cor = layerCor(stack, "pearson", na.rm = TRUE)
cor$pearson # 0.07553126  

```

## Check correlation

```{r check corr}

TWI_r = as.numeric(values(raster::raster(TWI_proj)))
NDVI_r = as.numeric(values(raster::raster(NDVI)))

# Calculate Pearson's correlation coefficient
correlation <- cor(TWI_r, NDVI_r, use = "pairwise.complete.obs", method = "pearson")

correlation # 0.09516853

```

