---
title: "SHAP figure (5 most imp)"
author: "Stenka Vulova"
date: '2024-04-22'
output: html_document
---

This is an example with the year 2018.

In the final figure, please use SHAP values averaged across all years (Katharina).

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
library(rgdal)
library(lubridate)
library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(terra)

library(ggspatial)
#library(geobuffer)
library(rgeos)

library(tools)
library(tibble)

```


# Import mean SHAP dataframes of the five most important predictors 

```{r import mean SHAP dfs }

tic("load mean SHAP df of the 5 most important predictors")

shap_df_path_tcd = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/TCD_mean_SHAP_df.csv"
shap_df_tcd <- data.table::fread(shap_df_path_tcd, select = c(1,2,8), header = TRUE)

shap_df_path_broadleaf = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/broadleaf_perc_mean_SHAP_df.csv"
shap_df_broadleaf <- data.table::fread(shap_df_path_broadleaf, select = c(1,2,8), header = TRUE)

shap_df_path_canopy = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/canopyheight_mean_SHAP_df.csv"
shap_df_canopy <- data.table::fread(shap_df_path_canopy, select = c(1,2,8), header = TRUE)

shap_df_path_spei = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/SPEI_sameyear_mean_SHAP_df.csv"
shap_df_spei <- data.table::fread(shap_df_path_spei, select = c(1,2,8), header = TRUE)

shap_df_path_et = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240423_mean_SHAP_dfs/ET_1year_mean_SHAP_df.csv"
shap_df_et <- data.table::fread(shap_df_path_et, select = c(1,2,8), header = TRUE)

toc()

tic.clearlog()

glimpse(shap_df_canopy)

# load mean SHAP df of the 5 most important predictors: 172.4 sec elapsed
# glimpse(shap_df_canopy)
# Rows: 10,439,063
# Columns: 3
# First two columns are the coordinates, third column is the mean_SHAP value for each coordinate 

```

```{r rename the colnames}

colnames(shap_df_tcd)[3] <- "TCD"
colnames(shap_df_broadleaf)[3] <- "broadleaf_perc"
colnames(shap_df_canopy)[3] <- "canopyheight"
colnames(shap_df_spei)[3] <- "SPEI_sameyear"
colnames(shap_df_et)[3] <- "ET_1year"

glimpse(shap_df_canopy)
# x 
# y
# canopyheight
# It worked!

```

```{r merge Top 5 Predictor dataframes to a single dataframe}

combined_df <- cbind(shap_df_tcd[, c("TCD")], shap_df_broadleaf[, c("broadleaf_perc")], 
                     shap_df_canopy[, c("canopyheight")], shap_df_spei[, c("SPEI_sameyear")], 
                     shap_df_et[, c("ET_1year")])

glimpse(combined_df)
# looks fine
# I will add the coordinate columns later, because otherwise, the next step won't work. 

```

# Most important predictor 

The most important of these 5 predictors is calculated per row (coordinate).
The highest absolute SHAP value means most important predictor for that coordinate. 

```{r most important predictor }

# Find the column index of the most important predictor for each row
most_important_col <- max.col(m = abs(combined_df), ties.method = "first")

# Get the column names corresponding to the most important predictor
most_important_predictor <- colnames(combined_df)[most_important_col]

# Create a list of the most important predictor for each row
most_important_list <- as.list(most_important_predictor)

# Print the list
head(most_important_list)
# seems fine

```

# Coordinates dataframe 

I am adding the coordinates from one of the shap_dfs I already loaded. 

```{r load coords }

# first two columns are x and y 

coords_df <- shap_df_tcd[, 1:2]

glimpse(coords_df)

# Rows: 10,439,063
# Columns: 2
# x
# y

```

# Make into one df 

```{r new df with coords }

imp_df_coords = cbind(coords_df, most_important_predictor)

glimpse(imp_df_coords)

# Rows: 10,439,063
# Columns: 3
# x
# y
# most_important_predictor

```

# levels of predictors

```{r convert to raster }

imp_df_coords$most_important_predictor = as.factor(imp_df_coords$most_important_predictor)

levels(imp_df_coords$most_important_predictor)

# [1] "broadleaf_perc" "canopyheight"   "ET_1year"       "SPEI_sameyear"  "TCD"        


levels_imp = data.frame(ID=1:5, most_important_predictor = c("broadleaf_perc", "canopyheight", "ET_1year", "SPEI_sameyear", "TCD" ))

levels_imp

# id
#1	broadleaf_perc			
#2	canopyheight			
#3	ET_1year			
#4	SPEI_sameyear			
#5	TCD

# add the id 1-5
imp_df_coords = dplyr::full_join(imp_df_coords, levels_imp)

```

# Convert to raster

```{r rasterize it }

shap_df_plot = imp_df_coords
shap_df_plot$most_important_predictor = NULL

# make a raster
mostimp_r = terra::rast(shap_df_plot, type = "xyz", crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

# the reason I converted to lonlat is to include the scalebar (it didnt work with the other projection)
# interpolate w/ nearest neighbor so that integers remain as they are
mostimp_lonlat = terra::project(x = mostimp_r, y = "+proj=longlat +datum=WGS84 +no_defs", method = "near")


levels(mostimp_lonlat) = levels_imp

terra::plot(mostimp_lonlat, col = c("#9467bd", "#d62728", "#ff7f0e", "#1f77b4", "#2ca02c"))
      
```

```{r Save raster and csv file}

# Save raster
writeRaster(mostimp_lonlat, "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240430_mostimp_SHAP_Fig6/mostimp_predictor_raster.tif", overwrite = TRUE)

# Load raster and check
raster <- rast("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240430_mostimp_SHAP_Fig6/mostimp_predictor_raster.tif")
plot(raster) 
# looks fine!



# Save mostimp predictor table
write.table(imp_df_coords, file = "D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240430_mostimp_SHAP_Fig6/mostimp_predictor_df.csv", row.names = FALSE, col.names = TRUE)

# Load mostimp predictor table and check
test <- data.table::fread("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240430_mostimp_SHAP_Fig6/mostimp_predictor_df.csv", header = TRUE)
glimpse(test)
# looks fine!

```

# Import Brandenburg and Berlin border

Keep Berlin in the plots :) 

```{r import border }

BB_border = readOGR(dsn = "D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border",
                        layer = "Brandenburg_border")

crs(BB_border)
# CRS arguments: +proj=longlat +datum=WGS84 +no_defs 

plot(BB_border)

crs <- crs(mostimp_lonlat)

# +proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs
BB_LAEA = sp::spTransform(BB_border, CRSobj = crs)
BB_border = BB_LAEA

```
```{r load cblind.plot() package and have a look at the colours}

library(RColorBrewer)
RColorBrewer::brewer.pal.info

# Load the "Paired" palette
paired_colors <- brewer.pal(12, "Paired")

# I ended up not using these colours, but different ones. 

```


# Plot 

```{r plot most imp }

png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240430_mostimp_SHAP_Fig6/20240430_shap_mostimp_map.png", width = 8, height =7, units = "in", res = 300)

# Set outer margins to accommodate legend
par(oma = c(0, 0, 0, 3)) # I added this, because otherwise the legend was cut off. 

raster::plot(mostimp_lonlat, col = c("orange", "magenta", "darkblue", "cyan", "#2ca02c"),
             legend.args = list(text = "Most contributing predictor", line = 0.5, cex = 1, adj = 0.25),
             legend = TRUE, axes = FALSE, box = FALSE)

#plot(mostimp_lonlat, col = c("green3", "magenta3", "gold2", "dodgerblue", "forestgreen"),
#     legend.only = TRUE, horizontal = FALSE , smallplot = c(0.87, 0.9, 0.17, 0.86),
#     axis.args = list(cex.axis = 1.4), 
#	legend.args = list(text = "Most contributing predictor", line = 0.5, cex = 1, adj = 0.25))

# add scalebar 
raster::scalebar(d = 50, # distance in km,
                 xy = c(11.5, 51.5),
                 type = "bar",
                 divs = 2,
                 below = "km",
                 lonlat = TRUE,
                 adj = c(0.5, -1.15),
                 cex = 1.2)

raster::plot(BB_border, add = TRUE)


dev.off()

```
