---
title: "NDMI_anomaly mapping"
author: "Katharina Horn"
date: '2024-05-10'
output: html_document
---

# Libraries 

```{r libs, include = FALSE }

library(sp)
library(raster)
#library(rgdal)
library(lubridate)
#library(maptools)
library(ggplot2)
library(tidyr)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(reshape2)
library(scales)
library(readr)
#library(rgeos)

library(grid)
library(spatstat)
library(sf)

library(zoo)
library(tictoc) # benchmarking

library(ggpubr)  # plotting multiple maps together
library(patchwork) # to use plot_spacer()
library(cowplot) # to use plot_grid()

library(terra)

library(ggspatial)
#library(geobuffer)

library(tools)
library(tibble)

library(scales)

```


```{r Load NDMI anomaly rasters}

NDMI <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Dynamic_data_proj/Annual/NDMI_annual_2018_to_2022.tif")

NDMI_2018 <- NDMI[[1]]
NDMI_2019 <- NDMI[[2]]
NDMI_2020 <- NDMI[[3]]
NDMI_2021 <- NDMI[[4]]
NDMI_2022 <- NDMI[[5]]

```

# Reproject Berlin-Brandenburg border

No need to redo this, since I have already run this part of code once. 

```{r Reproject Berlin-Brandenburg border}

bb <- terra::vect("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border.shp")

bb_proj <- terra::project(bb, NDMI_2018)
bb_proj
plot(bb_proj)

```

```{r}
# Load as sf object
bb_st <- st_read("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/Brandenburg Border/Brandenburg_border_3035.shp") # for mapping

```


```{r Load DEM layer to use for excluding Berlin NDMI values}

dem <- terra::rast("D:/Nextcloud/Documents/CliWaC/04_PROCESSED DATA/20230926_data_proj/Static_data_proj/DEM_proj.tif")

dem_proj <- terra::project(dem, NDMI)

```


```{r Mask NDMI layers to dem to exclude values within Berlin}

NDMI_2018_mask <- terra::crop(NDMI_2018, dem, mask = TRUE)
NDMI_2019_mask <- terra::crop(NDMI_2019, dem, mask = TRUE)
NDMI_2020_mask <- terra::crop(NDMI_2020, dem, mask = TRUE)
NDMI_2021_mask <- terra::crop(NDMI_2021, dem, mask = TRUE)
NDMI_2022_mask <- terra::crop(NDMI_2022, dem, mask = TRUE)

```


```{r crop NDMI layers to BB border}

NDMI_2018_crop <- raster::crop(NDMI_2018_mask, bb_proj)
NDMI_2019_crop <- raster::crop(NDMI_2019_mask, bb_proj)
NDMI_2020_crop <- raster::crop(NDMI_2020_mask, bb_proj)
NDMI_2021_crop <- raster::crop(NDMI_2021_mask, bb_proj)
NDMI_2022_crop <- raster::crop(NDMI_2022_mask, bb_proj)

```


```{r convert rasters into dataframes}

# Convert rasters into data frames
NDMI_2018_df <- as.data.frame(NDMI_2018_crop, xy = TRUE)
NDMI_2019_df <- as.data.frame(NDMI_2019_crop, xy = TRUE)
NDMI_2020_df <- as.data.frame(NDMI_2020_crop, xy = TRUE)
NDMI_2021_df <- as.data.frame(NDMI_2021_crop, xy = TRUE)
NDMI_2022_df <- as.data.frame(NDMI_2022_crop, xy = TRUE)

```

```{r Add additional column for the response variable}

NDMI_2018_df$response_var <- ifelse(NDMI_2018_df$`2018` >= -10, 0, 1)
NDMI_2019_df$response_var <- ifelse(NDMI_2019_df$`2019` >= -10, 0, 1)
NDMI_2020_df$response_var <- ifelse(NDMI_2020_df$`2020` >= -10, 0, 1)
NDMI_2021_df$response_var <- ifelse(NDMI_2021_df$`2021` >= -10, 0, 1)
NDMI_2022_df$response_var <- ifelse(NDMI_2022_df$`2022` >= -10, 0, 1)

# 0 means "no change", all values above -10 are assigned as "no change"
# 1 means "damaged", all values below -10 are assigned as "damaged" 
```

```{r Convert response_var into a factor with consistent levels in all data frames}

NDMI_2018_df$response_var <- factor(NDMI_2018_df$response_var, levels = c(0, 1))
NDMI_2019_df$response_var <- factor(NDMI_2019_df$response_var, levels = c(0, 1))
NDMI_2020_df$response_var <- factor(NDMI_2020_df$response_var, levels = c(0, 1))
NDMI_2021_df$response_var <- factor(NDMI_2021_df$response_var, levels = c(0, 1))
NDMI_2022_df$response_var <- factor(NDMI_2022_df$response_var, levels = c(0, 1))

```

```{r Define color scale}

colors <- c("0" = "#39b5ff","1" = "#F82602")

#colors <- colorRampPalette(c("#020d47", "#002ae7", "#39b5ff", "#00ff9b", "#fcf599", "#ff980c", "#F36802", "#F82602", "#7F0000"))(255)
#colors_rev <- c("#7F0000", "#F82602", "#F36802", "#ff980c", "#fcf599", "#00ff9b", "#39b5ff", "#002ae7", "#020d47")

```

# Plot NDMI for each year

I decided not to plot the north arrow, because it does not look good on the final map. I will add it with Inkscape afterwards. 

```{r Plot NDMI for all years}

# Plot NDMI_anomaly 2018
map_ndmi_2018 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = factor(response_var)), data = NDMI_2018_df) +       
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_manual(values = colors, name = "Response Variable", labels = c("no change", "damaged")) +      # set legend colours
  labs(x = NULL, y = NULL)  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2019
map_ndmi_2019 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = factor(response_var)), data = NDMI_2019_df) +       
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_manual(values = colors, name = "Response Variable", labels = c("no change", "damaged")) +      # set legend colours
  labs(x = NULL, y = NULL)  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2020
map_ndmi_2020 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = factor(response_var)), data = NDMI_2020_df) +       
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_manual(values = colors, name = "Response Variable", labels = c("no change", "damaged")) +      # set legend colours
  labs(x = NULL, y = NULL)  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2021
map_ndmi_2021 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = factor(response_var)), data = NDMI_2021_df) +       
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_manual(values = colors, name = "Response Variable", labels = c("no change", "damaged")) +      # set legend colours
  labs(x = NULL, y = NULL)  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

# Plot NDMI_anomaly 2022
map_ndmi_2022 <- ggplot() +
  geom_tile(aes(x=x, y=y, fill = factor(response_var)), data = NDMI_2022_df) +       
  geom_sf(bb_st, mapping = aes(), color = 'black', fill = NA) +                       
  scale_fill_manual(values = colors, name = "Response Variable", labels = c("no change", "damaged")) +      # set legend colours
  labs(x = NULL, y = NULL)  +
  scale_x_continuous(breaks = seq(11.0, 15.0, by = 1)) +
  scale_y_continuous(breaks = seq(51.0, 54, by = 1)) +
  annotation_scale(style = "bar", location = "bl") +                     # add scale bar
  theme(legend.position = "right") +
  theme_void()                                                             # no theme

```

# Check results before plotting everything.

```{r Check resulting plot for 2018}

map_ndmi_2018

# png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240510_NDMI_anomaly/NDMI_2018.png", width = 17.5, height = 12, units = "cm", res = 300)
# map_ndmi_2018
# dev.off()

```

# Get legend of one of the plots

I can extract a legend from one of the plots since all of them are based on the same min() and max() values
for each year. 

```{r Get legend of one of the plots to use as joint legend for all plots}

leg <- get_legend(map_ndmi_2018)

# Convert to a ggplot object
leg <- as_ggplot(leg)
leg

```

```{r Arrange resulting plots}

png("D:/Nextcloud/Documents/CliWaC/05_RESULTS/20240510_NDMI_anomaly/Response_variable_figure.png", width = 17.5, height = 12, units = "cm", res = 300)

response_vars <- ggarrange(map_ndmi_2018, map_ndmi_2019, map_ndmi_2020, map_ndmi_2021, map_ndmi_2022, leg,
          labels = c("2018", "2019", "2020", "2021", "2022", " "), 
          ncol = 3, nrow = 2, legend = "none")
response_vars

dev.off()

```
